---
title: "Hands-on Exercise 1"
author: "Hou Tao"
---

#### Getting Started

##### Install and Launching R packages

The code chunk below use p_load of pacman package to check if tidyverse packages are installed in te environment, it will luanch it into R if it has been installed.

```{r}

pacman::p_load(tidyverse)
pacman::p_load(ggplot2)

```

##### Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv", show_col_types = FALSE)
```

##### View Math Result using R hist

```{r}
hist(exam_data$MATHS,
     main = "Histogram of Maths Exams Result",
     xlab = "Score",
     xlim = c(0, 100),
     ylab = "Count",
     ylim = c(0, 80),
     col = "lightblue",
     freq = TRUE)
```

##### View Math Result using ggplot

```{r}
ggplot(data = exam_data, aes(x = MATHS)) + 
  geom_histogram(bins = 10, 
                 boundary = 100, 
                 color = "black",
                 fill = "lightblue") +
  labs(title = "Histogram of Exams Result", 
       subtitle = "Maths",
       x = "Score",
       y = "Count")
```

##### View Math Result using ggplot2 dotplot

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_dotplot(binwidth = 2.5, dotsize = 0.5) +
  scale_y_continuous(NULL, breaks = NULL)
  
```

##### View Math Result using ggplot2 histogram

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 color = "black",
                 fill = "lightblue")

```

##### View Math Result by Gender using ggplot2 histogram

```{r}
ggplot(data = exam_data,
       aes(x = MATHS, 
           fill = GENDER)) +
  geom_histogram(bins = 20,
                 color = "grey30") 

```

##### View Math Result using ggplot density
```{r}

ggplot(data = exam_data,
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()

```


##### View Math Result using ggplot boxplot
```{r}

ggplot(data = exam_data,
       aes(y = MATHS, 
           x = GENDER)) +
  geom_boxplot()

```


##### View Math Result using ggplot notched boxplot
```{r}

ggplot(data = exam_data,
       aes(y = MATHS, 
           x = GENDER)) +
  geom_boxplot(notch = TRUE)

```


##### View Math Result using ggplot boxplot and point
```{r}

ggplot(data = exam_data,
       aes(y = MATHS, 
           x = GENDER)) +
  geom_boxplot() +
  geom_point(position = "jitter", size = 0.5)

```


##### View Math Result using ggplot violin plot

> 1. violin plot is a hybrid of a box plot and a kernel density plot shows peaks in the data. it can visualize the distribution of numerical data.
> 
> 2. violin plot depict summary statistics and the density of each variable.
> (source: https://mode.com/blog/violin-plot-examples)

```{r}
ggplot(data = exam_data,
       aes(y = MATHS, 
           x = GENDER)) +
  geom_violin()

```





































